CTA Proposal:\\SEGUE 1: An Unevolved Fossil Galaxy from the Early Universe

### 2021-03-03 Phys 311/CTA/Paper/paper.tex
first draft

When studying the early universe scientists commonly seek out stars with low metal content. Locked inside these stars
is information about the physical and chemical conditions that led to the formation of the star. The low metallicity of
these stars, specifically low [Fe/H] values and high [$\alpha$/Fe] values, indicate that the stars were formed in the
early universe. By comparing the quantity of $\alpha$-elements, iron, and hydrogen we can map the timescales of
nucleosynthesis to the timescales of different types of supernovae. The $\alpha$-elements that the authors are interested 
in are carbon, oxygen, neon, magnesium, silicon, sulfur, argon, and calcium.

The earliest supernovae events in the universe were core-collapse supernovae of Massive stars. Massive stars produce large
amounts of $\alpha$-elements during their stellar evolution and core-collapse explosion. Massive stars have short lifetimes
of less than 10 million years. Heavy metals, such as iron, appeared later in the astronomical timeline as a byproduct of
Type Ia supernovae. Type Ia supernovae originate from binary star systems containing a white dwarf stars. These binary
systems have much longer lifetimes, on the order of $10^8$ years, before collapsing into a supernova. These substantially
different timelines allow scientists to approximately date the relative timing of nucleosynthesis in early stars.

For decades scientists looked at low metal stars in the halo of the Milky Way to study the early universe. In the past decade
efforts such as the Sloan Digital Sky Survey (SDSS) have uncovered a plethora of very dim dwarf spheroidal (dSph) galaxies and
ultra faint dwarfs with total luminosities in the range of \(10^{5} L_{\odot} \lesssim L \lesssim 10^{7} L_{\odot}\) and
\(L \lesssim 10^5 L_{\odot}\), respectively. Within this pool of dSph and ultra faint dwarf galaxies scientists discovered
a small handful of ultra faint dwarf galaxies that showed extremely low [Fe/H] values. In this small handful one galaxy,
SEGUE 1 stood out from the rest. SEGUE 1 stood out since it's stars show increasing metallicity but do not exhibit decreasing
[$\alpha$/Fe] ratios. This would indicate that SEGUE 1 dates back to the very early universe predating Type Ia supernovae events.

@@

The authors of this paper made new spectrography measurements of six stars within SEGUE 1 using the Magellan/MIKE and Keck/HIRES
spectrographs focusing on the chemical enrichment process to better understand star formation in the very early universe. The six
stars that were chosen were the brightest stars in SEGUE 1 which allowed for high resolution spectrography. Using the spectrographs
the authors collected data on the abundance of iron (Fe I and Fe II), carbon, $\alpha$-elements, neutron-capture elements, and other 
lighter elements. The authors then compared this data to stars found in classical dSph galaxies and the Milky Way halo.
The parameters for the measurements of each of the stars are listed in table 1. Due to the dim nature of the six starts the signal to
noise ratio (S/N) was only moderate with values between 20 and 50

@@

With the high resolution spectrograph data the authors were able to extrapolate the abundances of the various elements inside the stars.

In the three lowest metallicity stars, out of the six observed stars, the abundance of iron (both Fe I and Fe II) indistinguishable
from the noise in the measurements. In the 3 stars with observable amounts of iron there was no correlation between increasing
metallicity and increasing iron abundance. This indicated that there was only one Type Ia supernova that enriched SEGUE. This
contrasts the trend found in galaxies from later where there would be a clear trend in increasing metallicity and increasing
iron abundance as they evolve alongside energetic supernovae events. Other indications of the age of SEGUE include enhanced
[$\alpha$/Fe] values found in all it's stars that are consistent with their birth cloud, and the low abundance of neutron-capture
elements. The consistent [$\alpha$/Fe] values in SEGUE 1 stars suggest that there were no energetic supernovae events that enriched
the birth gas cloud during the star formation events. The low abundance of neutron-capture elements in SEGUE's non-binary stars,
which total to less than \(\sim 10^{-7} M_{\odot}\) for each element, indicate that there were no neutron-capture element producing
process, which produce \(\sim 10^{-4} M_{\odot}\) of neutron-capture elements per event, that occurred during the star formation phase.

>>>>

When studying the early universe scientists commonly seek out stars that were formed early on in the universe.
Locked inside these stars are clues about the physical and chemical conditions that led to the formation of the star.
With this information scientists can better understand physical and chemical conditions in the early universe as a whole.
One way to identify these ancient stars is by their relatively low heavy metal $\left(Z \geq 24\right)$ content compared to
lighter elements, such as $\alpha$-process elements and hydrogen.

Lighter $\alpha$-elements are synthesized within massive stars and are released into the interstellar medium (ISM) once these
massive stars undergo core-collapse supernova upon running out of helium. The lifetime of these massive stars, from their
nucleosynthesis to their core collapse supernova event takes place in a very short time frame, often less than $10^{6}$ yrs.
Heavier elements, the paper specifically measures iron (FeI and FeII), strontium, and barium abundance, are synthesized
and released later in the universe timeline. Iron is mainly synthesized during Type Ia supernovae events. These events occur
when a white dwarf in a binary star system reaches the critical Chandrasekhar limit and explodes into a supernova event.
These white dwarf systems have much longer lifetimes of around $10^{8}$ yrs.
Strontium and Barium, unlike the $\alpha$-elements and iron do not form via normal fusion but only during neutron capture
nucleosynthesis. Neutron capture nucleosynthesis happens during r-process supernovae and during the nucleosynthesis of
population III supermassive stars. During each of these neutron capture nucleosynthesis events approximately
$10^{-4} M_\odot$ of neutron capture elements are released into the ISM.

The low metallicity of early stars, specifically stars with high [$\alpha$/Fe] and low [Fe/H] values, indicate that their
nucleosynthesis occurred during a period after the first round of core-collapse supernovae but before large numbers of
Type Ia supernovae events. For decades scientists looked early stars in the halo of the Milky Way to study the early universe.
These Milky Way halo stars have [Fe/H] values of $< -1.0$.

In the past decade efforts such as the Sloan Digital Sky Survey (SDSS) have uncovered a plethora of very dim dwarf spheroidal
(dSph) galaxies and ultra faint dwarfs. These dSph and ultra faint dwarfs have total luminosities in the range of
\(10^{5} L_{\odot} \lesssim L \lesssim 10^{7} L_{\odot}\) and \(L \lesssim 10^5 L_{\odot}\), respectively, which contain
stars with [Fe/H] values of $\sim -2.5$. With the use of medium resolution spectroscopy scientists recently uncovered a
moderate population of super low metal stars with [Fe/H] values $< -3.0$ within ultra fain dwarfs. Within the same systems
scientists did not find any stars with [Fe/H] $> -1.0$. In this population one galaxy, SEGUE 1, stood out from the rest. 
SEGUE 1's population of stars show increasing metallicity, in the form of $\alpha$-elements, but do not exhibit decreasing
[$\alpha$/Fe] ratios. 
This would indicate that the star formation phase in SEGUE 1 occurred before most, if not all, Type Ia supernovae events. This
also indicates that there was likely only one star formation phase within the galaxy.

@@

The authors of this paper made new spectrography measurements of six stars within SEGUE 1 using the Magellan/MIKE and Keck/HIRES
spectrographs focusing on the chemical enrichment process to better understand star formation in the very early universe. The six
stars that were chosen were the brightest stars in SEGUE 1 which allowed for high resolution spectrography. Using the spectrographs
the authors collected data on the abundance of iron (Fe I and Fe II), carbon, $\alpha$-elements, and neutron-capture elements. The authors then compared this data to stars found in classical dSph galaxies and the Milky Way halo.
The parameters for the measurements of each of the stars are listed in table 1. Due to the dim nature of the six starts the signal to
noise ratio (S/N) was only moderate with values between 20 and 50

@@

With the high resolution spectrograph data the authors were able to extrapolate the abundances of the various elements inside the stars.
In figure 1 the authors plotted the spectra of the 6 stars compared to two other known stars, the ultra low metal $CD -38^{\circ}245$
and the very bright, and much younger, Arcturus.

In the three lowest metallicity stars (The top 3 blue lines) the abundance of iron (both Fe I and Fe II) is mostly indistinguishable
from the noise in the measurements. In the 3 stars with observable amounts of iron there was no correlation between increasing
iron abundance and increasing metallicity, indicated by the appearance of more spectral lines. This indicated that there was at most one or maybe two Type Ia supernova that enriched SEGUE. This
contrasts the trend found in galaxies from later where there would be a clear trend in increasing metallicity and increasing
iron abundance as they evolve alongside energetic supernovae events. Other indications of the age of SEGUE include enhanced
[$\alpha$/Fe] values found in all it's stars that are consistent with their birth cloud, and the low abundance of neutron-capture
elements. The consistent [$\alpha$/Fe] values in SEGUE 1 stars suggest that there were no energetic supernovae events that enriched
the birth gas cloud during the star formation events. The low abundance of neutron-capture elements in SEGUE's non-binary stars,
which total to less than \(\sim 10^{-7} M_{\odot}\) for each element, indicate that there were no neutron-capture element producing
process, which produce \(\sim 10^{-4} M_{\odot}\) of neutron-capture elements per event, that occurred during the star formation phase.
